Exploring probability distributions for bivariate temporal granularities

### 2019-08-12 main.Rmd
changing structure in Section 2 to make it flow better

These calendar categorisations can be linear, circular or aperiodic. The calendar categorisations as "linear" if they respect the linear progression of time. We call these **linear time granularities**. Examples include hours, days, weeks and months. **Circular time granularities** accommodate periodicities in time such as hour of the day, and day of the week. **Aperiodic circular granularities** are neither linear nor circular, such as day of the month or public holidays.

@@

*These definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality.*

## Formal conceptualisation of calendar categorisations{#cirgran-def}

@@

It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year. Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.

@@

<!-- Formal definition of aperiodic circular granularity -->

@@

Note that most of the formulas are approximations only due to unequal month and year lengths, and due to the fact that there are not an integer number of weeks per month or weeks per year.

 <!-- Why we need single order up and multiple order up granularities -->
<!-- A time series which is characterised by being composed of repeating cycles. -->
Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data. Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series. To consider the exhaustive set of temporal regularities that might exist ein the data, we can categorise the set of temporal granularities to single or multiple order up granularities.

## Single-order-up and multiple-order-up granularities

@@

# Computation of multiple order-up granularities from single order-up granularities

@@

## Periodic single-order-up granularities

@@

## Mixed single-order-up granularities - circular or aperiodic

@@

### Computing single-order-up granularities:

>>>>

These calendar categorisations can be linear, circular or aperiodic. The calendar categorisations as "linear" if they respect the linear progression of time. We call these **linear time granularities**. Examples include hours, days, weeks and months. **Circular time granularities** accommodate periodicities in time such as hour of the day, and day of the week. **Aperiodic time granularities** are neither linear nor circular, such as day of the month or public holidays.

@@

The definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality. Hence, there is a need to define circular time granularities in a different approach.

## Circular time granularities{#circular-gran-def}

Suppose we have a tsibble with a time index in one column and keys and variables in other columns. A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index. A linear granularity is a mapping of row numbers to subsets of the time domain. For example, if the time index is days, then a linear granularity might be weeks, months or years. What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.

@@

<!-- It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year. Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.  -->

We want to use modular arithmetic to define circular granularity. Hence, we start with the definition of equivalence classes and then move on to define a circular granularity.

@@

## Aperiodic time granularities{#aperiodic-gran-def}
<!-- Formal definition of aperiodic circular granularity -->

@@

 <!-- Why we need single order up and multiple order up granularities -->
<!-- A time series which is characterised by being composed of repeating cycles. -->

Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data. Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series. To consider the exhaustive set of temporal regularities that might exist ein the data, we can categorise the set of temporal granularities to single or multiple order up granularities.

# Single-order-up and multiple-order-up granularities

@@

## Computation of single-order-up granularities

## Computation of multiple order-up granularities

@@

### Periodic single-order-up granularities

@@

### Mixed single-order-up granularities - circular or aperiodic

@@
